---
title: "2019 Kaggle Ml & DS Survey"
author: "Kuldeep Singh Chouhan"
date: "26/10/2019"
output: 
  html_document:
    number_section: false
    fig_caption: true
    toc: yes
    theme: cosmo
    highlight: tango
    code_folding: hide
---

# **1 Introduction**

Welcome to the third annual Mahcine Learning and Data Science Survey. In [*2019 Kaggle Machine Learning and Data Science Survey*](https://www.kaggle.com/c/kaggle-survey-2019) Kaggle conducted an industry-wide survey that presents a truly comprehensive view of the state of data science and machine learning. According to the description the survey has **19,717 usuable responses from 171 countries and territories** after cleaning.

In this Kernel I am going to tell the rich story about a particular subset of data science and machine learning community. My focus in this kernel is on the salary distribution amoung different groups.  

# **2 Loading data and libraries** {.tabset .tabset-fade .tabset-pills}

We will start our EDA by checking the dimensions of our dataset.

## Loading Libraries

```{r message=FALSE}
library(tidyverse)
library(ggthemes)
library(data.table)
library(grid)
library(gridExtra)
#library(knitr)
library(kableExtra)

```

## Load Data

```{r}

multiple <- fread("data2019/multiple_choice_responses.csv", drop = c(4, 8, 20, 35, 48, 50, 51, 52, 53, 54, 
                                                                      55, 69, 82, 95, 97, 110, 116, 131, 140, 148, 155, 
                                                                      168, 181, 194, 207, 220, 233, 246), encoding = "UTF-8", skip = 1) %>% mutate_if(is_character, as_factor)

questions <- fread("data2019/questions_only.csv", drop = 1)

```

## Questions Table

The table of questions from the 2019 Kaggle Data Science and Machine Learning Survey. There are total 34 questions asked in the survey. 

```{r}

questions <- gather(questions, sno , question)

questions[1:34,] %>%
  kable(format = "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
  column_spec(column = 1, width = "5em;") %>%
  column_spec(column = 2, width = "20em;") %>%
  scroll_box(width = "100%")

```

## Overview of datatable

The dataset **multiple** consists of 246 columns. The column name are questions asked dusring the survey so we are going to edit the column names as we will explore our data during the EDA. 

```{r}
multiple[1:5,] %>%
  kable(format = "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
        column_spec(column = c(1,2,3), width = "15em; min-width: 15em;") %>%
        column_spec(column = c(4), width = "10em; min-width: 10em;") %>%
        column_spec(column = 5:218, width = "25em; min-width: 25em;") %>%
        row_spec(1:5, align = "c") %>%
        scroll_box(width = "100%")

```

Since the dataset is huge. I have used [**kableExtra**](https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html) to generate the view of full datatable in the browser.

## Column Name

We need ot change the column names to suitable variable name. Since Column names are questions.

```{r}

vars <- c(gender = "What is your gender? - Selected Choice",
          country = "In which country do you currently reside?",
          age = "What is your age (# years)?",
          education = "What is the highest level of formal education that you have attained or plan to attain within the next 2 years?",
          role = "Select the title most similar to your current role (or most recent title if retired): - Selected Choice",
          salary = "What is your current yearly compensation (approximate $USD)?",
          size = "What is the size of the company where you are employed?",
          ml_work = "Does your current employer incorporate machine learning methods into their business?")

multiple <- multiple %>%
  rename(!!vars)

```


# **3 Gender Distribution**

Q. What is your gender? - Selected Choice"

```{r fig.height=5, fig.cap="Fig. 1"}

gen <- multiple %>%
  count(gender) %>%
  mutate(gender = fct_reorder(gender,n))

gen %>%
  mutate(percentage = str_c(as.character(round(n/sum(gen$n)*100,1)), "%")) %>%
  ggplot(aes(gender, n, fill = gender)) +
  geom_col() +
  geom_label(aes(label = percentage), position = position_dodge(width = 1)) +
  labs(x = "Gender", y = "Number of Respondents") +
  theme_hc() +
  theme(legend.position = "none") +
  ggtitle("Gender Distribution") +
  coord_flip()

```

There is **huge gender imbalance with 82% male and only 16% female.** This is very sad.


# **4 Age Distribution**

Q. What is your age (# years)? 

```{r fig.height=5, fig.cap="Fig. 2"}

multiple <- multiple %>%
  mutate(age = fct_relevel(age, var <- c("18-21", "22-24","25-29","30-34","35-39","40-44","45-49","50-54","55-59","60-69","70+")))

multiple %>%
  count(age) %>%
  mutate(percentage = str_c(as.character(round(n/sum(n)*100,1)), "%")) %>%
  ggplot(aes(age, n, fill = age)) +
  geom_col() +
  geom_label(aes(label = percentage)) +
  labs(x = "Age", y = "Number of respondents") +
  theme_hc() +
  theme(legend.position = "none", axis.text.x  = element_text(angle=90, hjust=1, vjust=0.9)) +
  ggtitle("Age Distribution")

```

Age distribution is right-skewed, with **the most common value being age 25-29 with 23% followed by 22-24 and 30-34 with 18% & 16% respectively.** Most of the respondents are from younger age group. I am amazed to see respondents at ages of 60+ & 70+. 

# **5 Country**

Q. In which country do you currently reside?

```{r fig.height=9, fig.cap="Fig. 3"}

multiple %>%
  count(country) %>%
  mutate(country = reorder(country, n, FUN = min), n) %>%
  ggplot(aes(country, n)) +
  geom_col(fill = "yellowgreen") +
  labs(x = "", y = "Number of respondents") + theme_hc() +
  theme(legend.position = "none") +
  ggtitle("Country of respondents") +
  coord_flip()

```

The **maximum number of respondents are from India followed by US** forming around **40%** of the total respondents followed by **Other** with *5%** respondents. **Others (country or territory received less than 50
respondents).**

# **6 Education**

Q. What is the highest level of formal education that you have attained or plan to attain within the next 2 years?

```{r fig.height=5, fig.cap="Fig. 4"}

multiple %>%
  mutate(education = fct_relabel(education, str_replace, "Some college/university study without earning a bachelor’s degree", "Some college/Uni study")) %>%
  mutate(education = fct_relabel(education, str_replace, "No formal education past high school", "High School")) %>%
  mutate(education = fct_relabel(education, str_replace, "I prefer not to answer", "No answer")) %>%
  mutate(education = na_if(education, "")) %>%
  filter(!is.na(education))%>%
  count(education) %>%
  mutate(education = reorder(education, n, FUN = min), n) %>%
  mutate(percentage = str_c(as.character(round(n/sum(n)*100,1)), "%")) %>%
  ggplot(aes(education, n, fill = education)) +
  geom_col() +
  labs(x = "Education", y = "Number of respondents") + theme_hc() +
  geom_label(aes(label = percentage)) +
  theme(legend.position = "none") +
  ggtitle("Education of respondents") +
  coord_flip()

```

**The highest level of education are Master's & Bachelor's degrees with 44% & 31% respectively.** The third major category is **Doctoral degree** with **14%** respondents.

# **7 Current Role**

Q. Select the title most similar to your current role (or most recent title if retired): - Selected Choice

```{r fig.height=5, fig.cap="Fig. 4"}
multiple %>%
  mutate(role = na_if(role, "")) %>%
  filter(!is.na(role))%>%
  count(role) %>%
  mutate(role = reorder(role, n, FUN = min), n) %>%
  mutate(percentage = str_c(as.character(round(n/sum(n)*100,1)), "%")) %>%
  ggplot(aes(role, n, fill = role)) +
  geom_col() +
  labs(x = "Role", y = "Number of respondents") + theme_hc() +
  geom_label(aes(label = percentage)) +
  theme(legend.position = "none") +
  ggtitle("Role of respondents") +
  coord_flip()
```

  * **21%** respondents are **Data scientist and Software Engineer making 14% of the total respondents.** 

* There are enthusiastic students at second place with around same number as **Data Scientist** forming up equivalent **21%.** Wow seems Data Sciencw is promising career :). Great!

* **Statisticians** are very rare with only **1.7%.**


# **8 Size of the company**

```{r fig.height=5, fig.cap="Fig. 4"}
multiple %>%
  mutate(size = na_if(size, "")) %>%
  filter(!is.na(size)) %>%
  count(size) %>%
  mutate(size = reorder(size, n, FUN = min), n) %>%
  mutate(percentage = str_c(as.character(round(n/sum(n)*100,1)), "%")) %>%
  ggplot(aes(size, n, fill = size)) +
  geom_col() +
  labs(x = "Size", y = "Number of respondents") + theme_hc() +
  geom_label(aes(label = percentage)) +
  theme(legend.position = "none") +
  ggtitle("Size of the company") +
  coord_flip()

```


